---
title: "Class 14 Vaccination Mini Project"
author: 'Ivy (PID: A15442572)'
date: "3/3/2022"
output: 
  pdf_document:
    latex_engine: xelatex
---

Here we downloaded the most recently reported data from "https://data.ca.gov/dataset/covid-19-vaccine-progress-dashboard-data-by-zip-code/resource/15702a90-aa5d-49bc-8621-a8129630725a"

```{r}
# Import vaccination data
vax <- read.csv("covid19vaccinesbyzipcode_test.csv")
head(vax)
```

>Q1. What column details the total number of people fully vaccinated? 

persons_fully_vaccinated

>Q2. What column details the Zip code tabulation area? 

zip_code_tabulation_area

>Q3. What is the earliest date in this dataset? 

2022-01-05

>Q4. What is the latest date in this dataset? 

```{r}
vax$as_of_date[nrow(vax)]
```


Let's get a quick overview of this csv
```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset? 

9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 

18338

>Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 

```{r}
round((18338/107604)*100,2)
```

>Q8. [Optional]: Why might this data be missing?

People might have received their vaccine outside the county of SD.


# Working with dates

One of the “character” columns of the data is as_of_date, which contains dates in the Year-Month-Day format.

```{r}
library(lubridate)
```
```{r}
age <- today() - ymd("1998-06-02")
age
```
What is that in years?
```{r}
time_length(age, "year")
```


```{r}
# Specify that we are using the year-month-day format
vax$as_of_date <- ymd(vax$as_of_date)
```

How many days have passed since the first vaccination report?
```{r}
today() - vax$as_of_date[1]
```
How many days between the first and most recent?
```{r}
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]
```
First I have to  make sure my covid vaccination date column is in lubridate format

> Q9. How many days have passed since the last update of the dataset? 

```{r}
today() - vax$as_of_date[nrow(vax)]
```
> Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)?

```{r}
unique_dates <- unique(vax$as_of_date)
length(unique_dates)
```
# Working with ZIP codes

```{r}
library(zipcodeR)
```

Where is the ZIP code located?
```{r}
geocode_zip('92037')
```

Find the distance between two ZIP codes.
```{r}
zip_distance('92037','92109')
```

What is the census?
```{r}
reverse_zipcode(c('92037', "92109") )
```


```{r}
# Pull data for all ZIP codes in the dataset
zipdata <- reverse_zipcode( vax$zip_code_tabulation_area )
```


# Focus on the San Diego Area
```{r}
library(dplyr)

sd <- filter(vax, county == "San Diego")

nrow(sd)
```


> Q11. How many distinct zip codes are listed for San Diego County?

```{r}
uzip <- unique(sd$zip_code_tabulation_area)
length(uzip)
```


> Q12. What San Diego County Zip code area has the largest 12 + Population in this dataset?

```{r}
sd.12 <- filter(vax, county == "San Diego")
large_12 <- which.max(sd.12$age12_plus_population)
sd$zip_code_tabulation_area[large_12]
```
> Q13. What is the overall average “Percent of Population Fully Vaccinated” value for all San Diego “County” as of “2022-03-01”? 

```{r}
full_vax_sd <- filter(vax, county == "San Diego",
                         as_of_date == "2022-03-01")
removed_na_list <- na.omit(full_vax_sd$percent_of_population_fully_vaccinated)
round(mean(removed_na_list)*100, 2)
```
```{r}
summary(full_vax_sd$percent_of_population_fully_vaccinated)
```


> Q14. Using either ggplot or base R graphics make a summary figure that shows the distribution of Percent of Population Fully Vaccinated values as of “2022-03-01”?

Make the histogram
```{r}
hist(full_vax_sd$percent_of_population_fully_vaccinated, breaks = 30)
```

```{r}
library(ggplot2)

ggplot(full_vax_sd) +
  aes(percent_of_population_fully_vaccinated) + 
  geom_histogram()
``` 

# Focus on UCSD/ La Jolla

```{r}
ucsd <- filter(sd, zip_code_tabulation_area=="92037")
ucsd[1,]$age5_plus_population
```
> Q15. Using ggplot make a graph of the vaccination rate time course for the 92037 ZIP code area:

```{r}
baseplot<- ggplot(ucsd) +
            aes(as_of_date,
                percent_of_population_fully_vaccinated) +
            geom_point() +
            geom_line(group=1) +
            ylim(c(0,1)) +
            labs(title = "Vaccination Rate in La Jolla, CA (92037)", 
                 x ="Date", 
                 y="Percent Vaccinated")
baseplot
```

# Comparing similar sized areas

```{r}
# Subset to all CA areas with a population as large as 92037
vax.36 <- filter(vax, age5_plus_population > 36144 &
                as_of_date == "2022-03-01")

head(vax.36)
```
> Q16. Calculate the mean “Percent of Population Fully Vaccinated” for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”. Add this as a straight horizontal line to your plot from above with the geom_hline() function?

Now add a line showing average vaccinate rate for all zip codes area with population as great as 92037
```{r}
hline.36 <- mean(vax.36$percent_of_population_fully_vaccinated, na.rm = T)

baseplot + geom_hline(yintercept =  hline.36,
                      linetype="dashed", color = "red")
```

> Q17. What is the 6 number summary (Min, 1st Qu., Median, Mean, 3rd Qu., and Max) of the “Percent of Population Fully Vaccinated” values for ZIP code areas with a population as large as 92037 (La Jolla) as_of_date “2022-03-01”?

```{r}
summary(vax.36$percent_of_population_fully_vaccinated)
```


> Q18. Using ggplot generate a histogram of this data.

```{r}
ggplot(vax.36) +
  aes(percent_of_population_fully_vaccinated) + 
  geom_histogram()+
  xlim(c(0,1))
```
> Q19. Is the 92109 and 92040 ZIP code areas above or below the average value you calculated for all these above?

```{r}
vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92040") %>%
  select(percent_of_population_fully_vaccinated)

vax %>% filter(as_of_date == "2022-03-01") %>%  
  filter(zip_code_tabulation_area=="92109") %>%
  select(percent_of_population_fully_vaccinated)
```
The means for both 92109 and 92040 are below the average we calculated.

> Q20. Finally make a time course plot of vaccination progress for all areas in the full dataset with a age5_plus_population > 36144.

```{r}
vax.36.all <- filter(vax, age5_plus_population > 36144)


ggplot(vax.36.all) +
  aes(as_of_date,
      percent_of_population_fully_vaccinated, 
      group=zip_code_tabulation_area) +
  geom_line(alpha=0.2, color= "deeppink4") +
  ylim(c(0,1)) +
  labs(x="Date", y= "Percent Vaccinated",
       title="Vaccination Rate Across California",
       subtitle="Only areas with populations above 36k") +
  geom_hline(yintercept = hline.36, linetype= "dashed", color = "lightseagreen")
```
> Q21. How do you feel about traveling for Spring Break and meeting for in-person class afterwards?

Quite hopeful since a large proportion has been vaccinated.

