---
title: 'Class 6: R Functions'
author: 'Ivy (PID: A15442572)'
date: "2/3/2022"
output:
  pdf_document: default
  html_document: default
---

Today we will explore R functions.

We will start with calculating a grade for these example students.

```{r}
# Example input vectors to start with
student1 <- c(100, 100, 100, 100, 100, 100, 100, 90)
student2 <- c(100, NA, 90, 90, 90, 90, 97, 80)
student3 <- c(90, NA, NA, NA, NA, NA, NA, NA)
```

We could use the mean() fxn to calculate and avg
```{r}
mean(student1)
mean(student2, na.rm = TRUE)

```

How does the is.na() fxn work? Let's try it on student2
```{r}
student2
is.na(student2)
```
We can use this result to get at our NA values (i.e.
the TRUE positions)
```{r}
student2[is.na(student2)]
```
```{r}
x<- 1:5
x
x[x>2] <-100
x
```

Put these two things together - mean() is.na() fxn

```{r}
student2[is.na(student2)]<-0
mean(student2)

student3[is.na(student3)]<-0
mean(student3)
```

Now we want to fnd the lowest score
```{r}
student1
which.min(student1)

```

The which.min() fxn tells us where the smallest value is.
```{r}
student1[-8]
mean(student1[-which.min(student1)])
```

Store the student grades into a variable.
Convert all NA scores to 0.
Remove the lowest score and find the mean from the new list.
```{r}
x<- student2
x[is.na(x)]<-0
x<- x[-which.min(x)]
mean(x)
```
We are close to our working code snippet that will be the 
body of our first function.
```{r}
# First set NA values to 0
x[is.na(x)]<-0
# Remove lowest score and calculate mean
mean( x[-which.min(x)])
```

Now we can turn this into out first function. We will call this function 'grade()'
All functions include
-a name
-a body
-the arguments
```{r}
grade <- function(x){
  # First set NA values to 0
  x[is.na(x)]<-0
  # Remove lowest score and calculate mean
  mean( x[-which.min(x)])
}
```

Let's try it out
```{r}
grade(student1)
grade(student2)
grade(student3)
```

> Q1. Write a function grade() to determine an overall grade from a vector of student homework assignment scores dropping the lowest single score. If a student misses a homework (i.e. has an NA value) this can be used as a score to be potentially dropped. Your final function should be adquately explained with code comments and be able to work on an example class gradebook such as this one in CSV format: “https://tinyurl.com/gradeinput”

```{r}
gradebook<- read.csv("https://tinyurl.com/gradeinput", row.names=1)
head(gradebook)
```
We can use the 'apply()' fxn to grade the whole class

```{r}
# apply( x, 1, FUN)
apply(gradebook, 1 ,grade)
scores <- apply(gradebook, 1 ,grade)
```

> Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook?

```{r}
which.max(scores)
```

>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall?

```{r}
avg.hw<- apply(gradebook, 2, median, na.rm=TRUE)
which.min(avg.hw)
```

```{r}
boxplot(gradebook)
```
>Q4. Optional Extension: From your analysis of the gradebook, which homework was most
predictive of overall score (i.e. highest correlation with average grade score)?

```{r}
cor(scores, gradebook$hw1)
```

Now apply over all columns (i.e. homeworks)
```{r}
apply(gradebook, 2, cor, scores)
```
```{r}
mask<- gradebook
mask[is.na(mask)]<-0
mask
apply(mask, 2, cor, scores)
```


>Q5. Make sure you save your Rmarkdown document and can click the “Knit” button to generate
a PDF format report without errors. Finally, submit your PDF to gradescope. 
