---
title: "Class 10- Introduction to Genome Informatics Lab"
author: 'Ivy (PID: A15442572)'
date: "2/17/2022"
output: pdf_document
---

# Examine 1000 Genome Data

Q5: What proportion of the Mexican Ancestry in Los Angeles sample population (MXL) are 
homozygous for the asthma associated SNP (G|G)?

```{r}
mxl <- read.csv("373531-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
table(mxl$Genotype..forward.strand.)/nrow(mxl)
```

```{r}
gbr <- read.csv("373522-SampleGenotypes-Homo_sapiens_Variation_Sample_rs8067378.csv")
table(gbr$Genotype..forward.strand.)/nrow(gbr)
```

## Expression by Genotype Analysis

I want to read my RNA-Seq Expression Results into R. This file is not a CSV but rather has fields separated by space.

```{r}
gnmtype <- read.table("rs8067378_ENSG00000172057.6.txt")
head(gnmtype)
```

First try at this question. Is the mean expression different based on genotypes?

```{r}
gnmtype$geno == "G/G"
```

```{r}
summary(gnmtype[gnmtype$geno == "G/G",]$exp)
```

Now check the other genotypes

```{r}
table(gnmtype$geno)
```

Summary of "A/G" and "A/A"

```{r}
summary(gnmtype[gnmtype$geno == "A/G",]$exp)
summary(gnmtype[gnmtype$geno == "A/A",]$exp)
```

## Makw a summary overview figure

Make a boxplot figure...

```{r}
library(ggplot2)

ggplot(gnmtype) + aes(geno, exp, fill = geno) + geom_boxplot(notch = TRUE)
```





